Jug: Software for Reproducible Computation In Python

### 2016-10-14 paper.tex
MIN Improve text

\title{Jug: Software for Reproducible Computation}
\author{Luis Pedro Coelho}

@@

In addition to being valuable to the wider community, which may wish to
reproduce or extend a published analysis, reproducible research practices allow
for better control over the project by the authors themselves. For example, if
necessary parameters to run a pipeline are kept separately from the code that
implements it (perhaps even in mind of the researcher), this leads to
error-prone analysis and opens up the possibility that when the results are to
be written up for publication, the researcher will no longer be able to even
completely describe the process that led to them.

@@

However, unlike make, jug is written in Python and tasks can consist of any
Python function and its arguments. This can include running external commands.
A jugfile (the file which specifies which tasks need to be run---named by
analogy to the makefiles of make) is a simple Python script with some added
notations (we will see just how little changes need to be made from a
conventional Python script).

@@

By default, jug looks for a file called \code{jugfile.py}, but any filename can
be used. Generically, we refer to the script being run as the jugfile.

@@

If run on a single processor, this will just run all of the tasks in order. It
is most interesting when it is run on multiple processors. There it executes
all of the tasks that can be executed in parallel.

Of course, the actual code is more complex than what is shown above,
particularly to make sure that the locking is performed correctly and that the
waiting step eventually times out (in order to handle the situation where
another process is hung).

@@

The filesystem can support all of the above operations if the backend is coded
correctly to avoid race conditions. This is the default backend, identified
simply by a directory name. Inside this directory, files named by a hexadecimal
representation of their hashes. Objects are saved using Python's \code{pickle}
module with zlib compression. As a special case, \code{numpy} arrays
\citep{numpy} are saved to disk directly. For functionality, there was no need
for this inelegant special case as \code{numpy} can be saved as a pickle dump,
but \code{numpy} arrays are a very common data type in scientific programming
and saving them directly allows for very fast saving and loading (they are
represented on disk as a header followed by the binary information they
contain).

@@

The output of computation can be obtained from jug in two ways: (1) One can
simply write a task that writes to a file in the desired format. (2) They can
be inspected interactively using the \emph{shell} subcommand. It is expected
that the first option will be used for the final version of the computation,
whilst the second one is most helpful during development.

@@

It is often the case that one improves a step in a computational pipeline and
wished to obtain the result of recomputing the pipeline while reusing as much
of the previously computed intermediate results. Before using Jug, I found this
to be a very error-prone operation: by not removing enough of the intermediate
files, it was easy to generate an irreproducible state where part of the
computations were made using one version of the code and part with another.

Therefore, jug adds some support for result invalidation. When a results from a
task are invalidated, all tasks which, directly or indirectly, depend on them
are also marked for deletion.

It is still necessary to manually invalidate tasks. An alternative, which I
considered, was to have the code which implements the task be taken into
account while computing the hash. While this would add another layer of
protection, it would still be possible to make mistakes. If the function
depended on other functions, especially if this was done in a dynamic way, it
could be hard to discover all dependencies (cf.\ sumatra~\footnote{Sumatra is a
tool to collect all dependencies that are necessary to run a particular
computation. Its webpage is located at
\url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even minute
refactoring of the code would lead to over eager recomputation. This could make
the developer wary of making improvements in their code, resulting in overall
worse code.

Therefore, I chose to force the user to explicitly invalidate his results,
while supporting automatic dependency discovery to make it easier for the user.
The recommendation is still that the user run the full pipeline from start to
finish once they are satisfied with the state of the code and certainly before
publication, but the pipeline development stage can be more agile.

\section{Methods}
\section{Example}
I will present an example from my research. Although it introduces some
unnecessary complexity, it has the advantage of being a real-life example of
the usage of jug.

@@

Jug has been available in the Python Package Index\footnote{The Python Package
Index, PyPI, is accessible at \url{http://pypi.python.org}.} since May~2009
through May~2013 and has been downloaded over 43,000~times (some of these
downloads may, of course, represent upgrades, and users which downloaded it
from some other source will not be counted).

@@

While it still possible to make errors (for example, by changing the code but
neglecting to invalidate the results), jug improves the pipeline development
experience and makes the programming researcher more productive and less
error-prone.

Jug was not designed to compete with complete map/reduce frameworks such as
Hadoop, which scale better to very large projects. However, those frameworks
have much higher costs in terms of development time (code must be written
especially for the framework) and overhead (starting up and shutting down jobs
can take much longer and using non-Java code has additional computational
costs). For scientists using small to medium-sized cluster (hundreds, not
thousands, of nodes), with a shared file-system (as opposed to the distributed
file-systems necessary for truly large amounts of data) jug might provide a
better trade-off than those higher-powered alternatives.

In the last few years, I have not undertaken any computational research project
without the use of jug to manage my computations. Additionally, I (and my
co-authors) have made the resulting results directories available\footnote{See
the links at
\url{http://murphylab.cbi.cmu.edu/software/software_from_papers.html}}. The
tool has become an indispensible part of my toolkit.

>>>>

\title{Jug: Software for Reproducible Computation In Python}
\author{Luis Pedro Coelho}

@@

In addition to being valuable to the wider community, which may wish to
reproduce or extend a published analysis, reproducible research practices allow
for better control over the project by the authors themselves. For example, if
necessary parameters to run a pipeline are kept separately from the code that
implements it (in some instance even in mind of the researcher), this leads to
error-prone analysis and opens up the possibility that when the results are to
be written up for publication, the researcher will no longer be able to even
completely describe the process that led to them.

@@

However, unlike make, jug is written in Python and tasks can consist of any
Python function and its arguments. This can include running external commands
and calling routines written in other languages as Python has many tools to
interface with the rest of the system.

A jugfile (the file which specifies which tasks need to be run---named by
analogy to the makefiles of make) is a simple Python script with some added
notations (we will see the only a few, small, changes are needed to transform a
conventional Python script into a Jugfile).

@@

By default, jug looks for a file called \code{jugfile.py}, but any filename can
be used. Generically, we refer to the script being run as the Jugfile.

@@

If run on a single processor, this will just run all of the tasks in order. It
is most interesting when it is run on multiple processors. Because of the lock
synchronisation, tasks can be run in parallel to take advantage of the multiple
cores.

The actual code is more complex than what is shown above, particularly to make
sure that the locking is performed correctly and that the waiting step
eventually times out (in order to handle the situation where another process is
hung).

@@

The filesystem can support all of the above operations if the backend is coded
correctly to avoid race conditions. This is the default backend, identified
simply by a directory name. Inside this directory, files named by a hexadecimal
representation of their hashes. Objects are saved using Python's \code{pickle}
module with zlib compression. As a special case, \code{numpy} arrays
\citep{numpy} are saved to disk directly. This special case was introduced as
\code{numpy} arrays are a very common data type in scientific programming and
saving them directly allows for very fast saving and loading (they are
represented on disk as a header followed by the binary information they
contain).

@@

The output of computation can be obtained from jug in three ways: (1) One can
simply write a task that writes the output to a file in the desired format. (2)
Outputs can be inspected interactively using the \emph{shell} subcommand. It is
expected that the first option will be used for the final version of the
computation, whilst the second one is most helpful during development. Finally,
(3) jug can be loaded as a library from a Python script and jug computation
outputs can serve as inputs for further computation.

@@

If a developer improves an intermediate step in a pipeline (e.g., fixes a bug)
and wishes to obtain new results of the computation, then all outputs from that
step and downstream must be recomputed, but results from upstream and unrelated
processes can be reused. Formally, in the task DAG, affected tasks and their
descendants must be recomputed. Without tool support, this can be a very
error-prone operation: by not removing enough of the intermediate files, it was
easy to generate an irreproducible state where part of the computations were
made using one version of the code and part with another.

Therefore, jug adds support for result invalidation. When a results from a task
are invalidated, all tasks which, directly or indirectly, depend on them are
also marked for deletion.

It is still necessary to manually invalidate tasks. An alternative, was to have
the code which implements the task be taken into account while computing the
hash. While this would add another layer of protection, it would still be
possible to make mistakes. If the function depended on other functions,
especially if this was done in a dynamic way, it could be hard to discover all
dependencies (cf.\ sumatra~\footnote{Sumatra is a tool to collect all
dependencies that are necessary to run a particular computation. Its webpage is
located at \url{http://neuralensemble.org/trac/sumatra}.}). Additionally, even
minute refactoring of the code would lead to over eager recomputation. This
could make the developer wary of making improvements in their code, resulting
in overall worse code.

Therefore, as a design choice, jug asks the user to explicitly invalidate their
results, while supporting automatic dependency discovery. The recommendation is
still that the user run the full pipeline from start to finish once they are
satisfied with the state of the code and certainly before publication, but the
pipeline development stage can be more agile.

\section{Methods}
\section{Example}

I will present an example from my research. Although it introduces some
unnecessary complexity, it has the advantage of being a real-life example of
the usage of jug.

@@

@@

Jug improves the pipeline development experience and makes the programming
researcher more productive and less error-prone. As it is based on Python, a
widely used scientific programming language, it can be used in a known
environment without a long learning process.

Jug was not designed to compete with large scale frameworks such as Hadoop or
Spark, which scale better to very large projects. However, those frameworks
have much higher costs in terms of development time (code must be written
especially for the framework) and overhead. For scientists using small to
medium-sized cluster (hundreds, not thousands, of nodes), with a shared
file-system (as opposed to the distributed file-systems necessary for truly
large amounts of data) jug provides a better trade-off than those
higher-powered alternatives in a familiar programming language.
